\Large\textsc{Regression And Resampling Techniques \\ FYS-STK4155 \\ Assignment One}

### 2019-09-24 assignment_one/article/main.tex
Working on article.

  \subsection{The bias-variance tradeoff}

  Since we are attempting to estimate a true distribution based on finite
  samples, there are certain statistical properties it is worthwile discussing.
  We follow the derivation from \autocite{mehtaHighbiasLowvarianceIntroduction2019}.  Recall
  \cref{as:general_linear} which states that our true data is generated from a
  noisy model
  \begin{equation}
    y = f(\x) + \varepsilon
  \end{equation}
  with \( \varepsilon \sim \N(0, \sigma^2) \).  We select a finite sample from
  the true data, which constitutes our finite data-set \( \data \).
  Furthermore, we have a model function \( f(\data, \beta) \) which depends on
  the selected data-set and the model parameters \( \beta \). Thus, our
  cost-function \( \cost \) will depend on the specific sample taken from the
  true distribution. It is of interest to examine the expected out-of-sample
  error \( \expect_{\data, \varepsilon}[\cost(\data, \beta)] \) of the model
  function over the sampled data set \( \data \) and the noise \( \varepsilon
  \). This can be decomposed into three distinct parts,
  \emph{\textcolor{Maroon}{variance}}, \emph{\textcolor{Blue}{bias}}, and
  \emph{\textcolor{Green}{noise}}:
  \begin{equation}
    \expect_{\data, \varepsilon}[\cost(\data, \beta)] = \frac{1}{n} \textcolor{Blue}{\sum_{i=1}^n \left(f(\x_i) - \expect_{\data}\left[\hat{y}_i\right]\right)^2}  + \frac{1}{n}\textcolor{Maroon}{\sum_{i=1}^n \expect_{\data}\big[\left(\hat{y}_i - \expect_{\data} [\hat{y}_i]\right)^2\big]} + \textcolor{Green} \sigma^2.
  \end{equation}
  See \cref{ap:derivation} for the full derivation. The bias-term measures the
  deviation of the expectation of the model function \(f(\data, \beta)\) from
  the true value \( f \), while the variance-term measures the fluctuations
  arising due to using different samples of the true data.

@@

  When the dimensionality of the matrix \( \X \) is high, the problem of
  near-collinear columns often arise, as mentioned earlier. Thus, the model
  parameters \( \beta_{\OLS} \) of the ordinary least squares regression cannot
  be determined. The propsed fix to this problem is to slightly perturb the
  matrix \( \X^T \X \) by adding small positive values along the diagonal to
  ensure its invertability. This is known as \emph{ridge regression} and yields
  a model parameter estimator
  \begin{equation}
    \beta_{\ridge}(\lambda) = (\X^T\X + \lambda\mat{I})^{-1}\X^T\y
  \end{equation}
  where \( \lambda \geq 0 \) is a hyper parameter. To see why the resulting
  modified matrix is invertible, one has to look at the singular values.  It
  turns out that the model parameters \( \beta_{\ridge}(\lambda) \) directly
  minimize the \emph{ridge loss function} defined as
  \begin{equation}
    \cost_{\ridge}(\data, \beta;  \lambda) \coloneqq \frac{1}{n} \sum_{i=1}^n
    (y_i - \hat{y}_i)^2 + \lambda \sum_{i=1}^p \beta_i^2 = \cost(\data, \beta)
    + \lambda \beta^T \beta,
  \end{equation}
  which we recognize as standard penalized least squares.

@@

>>>>

  \subsection{The bias-variance tradeoff} 
  \label{sub:bias_variance} 
  Since we are attempting to estimate a true distribution based on finite
  samples, there are certain statistical properties it is worthwile discussing.
  We follow the derivation from \autocite{mehtaHighbiasLowvarianceIntroduction2019}.  Recall
  \cref{as:general_linear} which states that our true data is generated from a
  noisy model
  \begin{equation}
    y = f(\x) + \varepsilon
  \end{equation}
  with \( \varepsilon \sim \N(0, \sigma^2) \).  We select a finite sample from
  the true data, which constitutes our finite data-set \( \data \).
  Furthermore, we have a model function \( f(\data, \beta) \) which depends on
  the selected data-set and the model parameters \( \beta \). Thus, our
  cost-function \( \cost \) will depend on the specific sample taken from the
  true distribution. It is of interest to examine the expected out-of-sample
  error \( \expect_{\data, \varepsilon}[\cost(\data, \beta)] \) of the model
  function over the sampled data set \( \data \) and the noise \( \varepsilon
  \). This can be decomposed into three distinct parts,
  \emph{\textcolor{Maroon}{variance}}, \emph{\textcolor{Blue}{bias}}, and
  \emph{\textcolor{Green}{noise}}:
  \begin{equation}
    \label{eq:bias_variance}
    \expect_{\data, \varepsilon}[\cost(\data, \beta)] = \frac{1}{n} \textcolor{Blue}{\sum_{i=1}^n \left(f(\x_i) - \expect_{\data}\left[\hat{y}_i\right]\right)^2}  + \frac{1}{n}\textcolor{Maroon}{\sum_{i=1}^n \expect_{\data}\big[\left(\hat{y}_i - \expect_{\data} [\hat{y}_i]\right)^2\big]} + \textcolor{Green} \sigma^2.
  \end{equation}
  See \cref{ap:derivation} for the full derivation. The bias-term measures the
  deviation of the expectation of the model function \(f(\data, \beta)\) from
  the true value \( f \), while the variance-term measures the fluctuations
  arising due to using different samples of the true data.

@@

  When the dimensionality of the matrix \( \X \) is high, the problem of
  near-collinear columns often arise, as mentioned earlier. Thus, the model
  parameters \( \beta_{\OLS} \) of the ordinary least squares regression cannot
  be determined. The propsed fix to this problem
  \cite{tikhonovNumericalMethodsSolution1995,hoerlRidgeRegressionBiased} is to
  slightly perturb the matrix \( \X^T \X \) by adding small positive values
  along the diagonal to ensure its invertability. This is known as \emph{ridge
  regression} and yields a model parameter estimator
  \begin{equation}
    \beta_{\ridge}(\lambda) = (\X^T\X + \lambda\mat{I})^{-1}\X^T\y
  \end{equation}
  where \( \lambda \geq 0 \) is a hyper parameter. To see why the resulting
  modified matrix is invertible, one has to look at the singular values.  It
  turns out that the model parameters \( \beta_{\ridge}(\lambda) \) directly
  minimize the \emph{ridge loss function} defined as
  \begin{equation}
    \cost_{\ridge}(\data, \beta;  \lambda) \coloneqq \frac{1}{n} \sum_{i=1}^n
    (y_i - \hat{y}_i)^2 + \lambda \sum_{i=1}^p \beta_i^2 = \cost(\data, \beta)
    + \lambda \beta^T \beta,
  \end{equation}
  which we recognize as standard penalized least squares.

@@

  We are also interested in computing the model bias and variance as discussed
  in \cref{sub:bias_variance}. Numerically, we estimate these quantities by
  replacing the expectation values in \cref{eq:bias_variance} by averages.

  \subsection{Cross-validation}

  In order to obtain more reasonable estimates for these statistical
  quantities, we perform so-called \emph{resampling} of the data. There are
  several approaches to resampling, with two of the more common ones being
  \emph{bootstrap} and \emph{k-fold cross-validation}. Without going into
  details, the bootstrap method generally performs better with a significant
  data size, with cross-validation also performing well on smaller data sets.
  With this in mind, we will discuss cross-validation here.
